---
title: "Journal 1"
#bibliography: references.bib
author: "Emily Kurtz"
output: 
  html_document:
    css: tweaks.css
    toc:  true
    toc_float: true
    number_sections: false
 
---



```{r, globalsettings, echo=FALSE, warning=FALSE, results='hide'}
library(knitr)

knitr::opts_chunk$set(echo = TRUE)
opts_chunk$set(tidy.opts=list(width.cutoff=100),tidy=TRUE, warning = FALSE, message = FALSE,comment = "#>", cache=TRUE, class.source=c("test"), class.output=c("test2"))
options(width = 100)
rgl::setupKnitr()



colorize <- function(x, color) {sprintf("<span style='color: %s;'>%s</span>", color, x) }

```

```{r klippy, echo=FALSE, include=TRUE}
klippy::klippy(position = c('top', 'right'))
#klippy::klippy(color = 'darkred')
#klippy::klippy(tooltip_message = 'Click to copy', tooltip_success = 'Done')
```

Last compiled on `r format(Sys.time(), '%B, %Y')`

<br>

----
  
White index code.
  

----




```{r}
# page 143 of Reardon and O'Sullivan, formula 13, shows the White spatial proximity index
# https://rdrr.io/cran/seg/man/isp.html

# sum(pi_m * P_mm / P_tt)
# basically just the distances for someone to their own people / to everyone else

# i am going to have someone input a square matrix that has all of the people of one group in the first chunk, second in second, etc. let's just start with two groups and a 10 by 10
# entries in cells are distances


# pi_m is the proportion of people in group m, matrix is the distances matrix for each individual to each other (symmetric matrix), and gpMember is an n by g matrix where n is the number of people and g is the number of groups


# okay think that works most of the way for two groups, so now i need to do for n groups
whiteIndex<-function(pi_m, matrix, gpMember){
    P_mm<-rep(0,ncol(gpMember))
    # calculate P_tt
    # this is just the mean of the pairwise distances
    P_tt <- (sum(rowSums(matrix))/2) / choose(2, nrow(matrix))
    for(i in 1:ncol(gpMember)){
      # calculate P_mm
      # this is just the mean of the pairwise distances for all pairs where both in gp i
      index <- which(gpMember[,i] == 1)
      P_mm[i] <- sum(rowSums(matrix[index,index]) / 2) / choose(2, sum(gpMember[,i]))
    }
    white <- sum(pi_m*P_mm/P_tt) #need to ensure pi_m is of length gp1size
    white
}
```

<br>

----
  
Answers to questions.
  

----



A. Read the following papers:

Reardon and O’Sullivan (2004)
Yao et al. (2019)
Hong, O’Sullivan, and Sadahiro (2014)
Tivadar (2019)

B. For this BIGSSS we will use GIS data from Statistics Netherlands, at the municipality, district (‘wijk’), (‘buurt’) and 100-by-100 meter grid level. What are the basic requirements of data necessary and sufficient to determine the level of spatial segregation in specific areas of the Netherlands? Make a list of requirements for non-spatial and spatial measures.

Non-spatial:

Evenness & exposure (census tract boundaries), group sizes, how they are distributed in space.

Spatial:

spatial proximity of neighborhoods (checkerboard issue), boundaries of units (MAUP issue), clustering, concentration, & centralization (location and size of census tract)

C. How would you theoretically want to define the spatial distance between voters/citizens in the Netherlands?

Measuring the distance using the Euclidean metric doesn't seem like the best idea. As a simple example, imagine two situations in a city that is organized along a perfect square grid. In the first, two friends live 10 north/south blocks and 10 east/west blocks away from each other. Their Euclidean distance is $\sqrt{200}$ blocks, but the friends have to walk 20 blocks to get to each other. In the second situation, the friends live on the same east/west road, but 20 blocks from each other. They still must walk 20 blocks, but the Euclidean distance is is also 20 blocks, which is much more than $\sqrt{200}$.

If the whole world were arranged on such a grid, then, I would advocate for measuring spatial distance using the L1 norm (Manhattan distance), but this is not the case. It would be good to measure distance as closely to accurate walking paths as possible. The L1 norm may, however, be a useful approximation in some cases.

D. Please formulate a precise definition of what you consider to be the relevant neighbourhood (or phrased otherwise ‘local environment’ or ‘social environment’) as input for the segregation measures.

E. Which theoretical article on residential segregation should we definitely all read for this summer school? Please provide a reference and motivate your answer.

Gimpel and various colleagues have some interesting theoretical work that is somewhat related. Specifically, some works discuss how rural and urban settings can differently impact political thinking. One paper that I recently read and liked is Gimpel and Reeves (forthcoming). As a sneak peek, here's a nice quote: “The greater familiarity among residents of low-density environments produces social pressures unlike the independence and unconventionality allowed by crowded nameless settings. Easy recognition means deviance from norms is highly visible.”

Since we are studying voting, a private act, instead of publicly stated political views, we may not see this pattern, but I thought it, and the work more broadly, would be interesting to consider going forward.

Gimpel, J. & Reeves, A. Political Geography, Political Behavior, and Public Opinion. (Forthcoming). 




----
  
Recreating the Reardon & O'Sullivan Visual Using iniworld
  

----


There are two dimensions: evenness (clustering) and exposure (isolation). Evenness refers to the degree to which the proportions of groups within a smaller subarea match their proportions in the overall population, and exposure refers to the degree to which members of one group encounter in their local environment members of each group. 




